Exploring Constraint Grammar as a SAT problem

### 2015-03-29 paper/nodalida2015.tex
more stuff
more text

\begin{verbatim}
"<la>"
        "el" det def f sg
        "lo" prn pro p3 f sg
"<casa>"
        "casa" n f sg
        "casar" v pri p3 sg
        "casar" v imp p2 sg
\end{verbatim}

The ambiguous passage can be either a noun phrase, \emph{la\texttt{:det} casa\texttt{:n}} 
`the house'  or a verb phrase \emph{la\texttt{:prn} casa\texttt{:v}} `(he/she) marries her'. 
We add the following rules:

@@

SAT solver 
performs a search, 
and starts building possible models that satisfy both constraints. 
In addition to the given constraints, we have default rules to emulate
the CG principles: an analysis is true if no rule affects it,
and at least one analysis for each word is true---the notion of ``last'' is not applicable.
With these constraints, we get two solutions: \texttt{det n} and \texttt{prn v}. 
The interaction of the rules regarding \emph{la}  disambiguates \emph{casa} 
for free, and ordering of the rules doesn't matter. 
%Without more context or additional rules we cannot fully disambiguate the passage,
%but unlike with standard CG, we get only legal combinations in one solution.

@@

@@

To replace the sequential order, we have experimented with different strategies:
\begin{itemize}
\item Maximise number of rules used---assume if someone has written a rule, it is worth applying. We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.
\item [\checkmark] Maximise number of instances when a rule is applied---assume that whenever we can apply a rule, we should do it. 
\item Maximise number of true analyses---Without this heuristic, both solutions to our example \emph{la casa} are as likely. With this heuristic,  the example would be disambiguated as \texttt{prn v}, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative.
% We want analyses to be true by default, but still disambiguate as much as possible.
%We have a rule to maximise the number of true analyses; then we take a solution that has the least. 
\item [\checkmark] Maximise rules, minimise number of true analyses---To combat the previous, we want each solution to contain as many true analyses as possible, to ensure that analyses untouched by rules are kept true. After having multiple solutions which all ensure , we pick the one that disambiguates the most.

\end{itemize}

@@

SAT is NP-complete, whereas the current implementations of CG are polynomial, 
but with advances in SAT solving techniques, 
the performance is much more feasible than in the previous works done in 90s--00s.
We used the open-source SAT solver MiniSat \cite{een04sat}.
%We do foreach sentence: foreach rule; 
%whereas VISL CG3 has foreach rule: foreach sentence.
%the size of the SAT problem is dependent of the sentence (???)

@@

\begin{table}
  \centering
  \begin{tabular}{|l|l|c|c|}
     \hline
	        \textbf{Text} & \textbf{\# words, rules} & \textbf{VISLCG3} &  \textbf{SATCG} \\ \hline
    \multirow{2}{*}{\textbf{Don Quijote}} & 186460 words  & 10.7s &  2m53.708s \\ 
		                                           & 261 rules & & \\ \hline
    \multirow{2}{*}{\textbf{Pride and}}  & 153386 words & 4.811s & 55.705s \\ 
                         {\textbf{Prejudice}}                   & 99 rules & \\ \hline
  \end{tabular}
  \caption{Comparing time between SATCG and VISL CG3.}
  \label{table:time}
\end{table}

@@

\paragraph{Removing rules}
\begin{table}
  \centering
  \begin{tabular}{|l|c|c|}
     \hline
    \textbf{\# rules} & \textbf{SATCG} & \textbf{VISLCG3} \\
    \hline 
	19 &  83.15 & 83.33 \\ \hline  % 19 out of 23 preselected by authors
	18 &  83.19 & 83.04 \\ \hline  % 18 out of 19
	17 &  83.201 & 83.03 \\ \hline   % 17 out of 18/19, no diff
	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with SATCG, chose to retain the rule that gave better results for VISLCG3.
	14 &  83.180 & 83.005 \\ \hline  
	13 &  & \\ \hline  
	12 &  & \\ \hline  
	11 &  & \\ \hline  
	10 &  & \\ \hline  
	2   & 80.23 & 79.97 \\ \hline
	1   & 76.69 & 76.62 \\ \hline
	0   & &  \\ \hline
  \end{tabular}
  \caption{}
  \label{table:results}
\end{table}
In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISLCG3.
Then we removed rules one by one, keeping rules with most significant gain

test the performance of the two implementations by working with a smaller grammar. We compare how much removing the rules affects the result, gold standard being for each implementation the result given by the full rule set.
\todo{Perform this experiment}

%5195:
%Select [[det],[detnt],[rel,aa]] (C (Exactly 1) (False,[[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]]))
%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly 1) (True,[[vblex],[vbmod],[vbhaver],[vbser],[adv],[prn,pro]])))
%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly (-1)) (True,[[pr]])))
%Select [[adv]] (C (Exactly 1) (True,[[],["sent"],[cm],[lpar],[rpar],["<;>"],[cnjcoo],[cnjsub],[cnjadv],[rel],[pr],[det],[detnt],[rel,aa]]))
%Select [[preadv]] (C (Exactly 1) (True,[[pp]]))
%Remove [[prs,p1],[pis,p1],[pii,p1],[cni,p1]] (C (Exactly (-1)) (True,[[n],[p3]]))
%Remove [[vblex],[vbmod],[vbhaver],[vbser]] (C (Exactly (-1)) (True,[[vblex]]))
%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (AND (C (Exactly 0) (True,[[pr]])) (C (Exactly 1) (True,[[inf]])))
%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (C (Exactly (-1)) (True,[[pr]]))
%6432:
%Select [[prn,pro],[prn,itg],[prn,tn]] (C (Exactly 1) (True,[[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]]))
%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly 1) (True,[[vblex],[vbmod],[vbhaver],[vbser],[adv],[prn,pro]])))
%Select [[prn]] (AND (C (Exactly 0) (True,[[adj]])) (C (Exactly (-1)) (True,[[pr]])))
%Select [[adv]] (C (Exactly 1) (True,[[],["sent"],[cm],[lpar],[rpar],["<;>"],[cnjcoo],[cnjsub],[cnjadv],[rel],[pr],[det],[detnt],[rel,aa]]))
%Select [[preadv]] (C (Exactly 1) (True,[[pp]]))
%Remove [[prs,p1],[pis,p1],[pii,p1],[cni,p1]] (C (Exactly (-1)) (True,[[n],[p3]]))
%Remove [[vblex],[vbmod],[vbhaver],[vbser]] (C (Exactly (-1)) (True,[[vblex]]))
%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (AND (C (Exactly 0) (True,[[pr]])) (C (Exactly 1) (True,[[inf]])))
%Remove [[pri],[prs],[imp],[pii],[cni],[pis],[ifi],[fti]] (C (Exactly (-1)) (True,[[pr]]))

>>>>

\begin{verbatim}
"<la>"
        "el" det def f sg
        "lo" prn p3 f sg
"<casa>"
        "casa" n f sg
        "casar" v pri p3 sg
        "casar" v imp p2 sg
\end{verbatim}

The ambiguous passage can be either a noun phrase, \emph{la\texttt{:det} casa\texttt{:n}} 
`the house'  or a verb phrase \emph{la\texttt{:prn} casa\texttt{:v pri p3}} `(he/she) marries her'. 
We add the following rules:

@@

SAT solver 
performs a search, 
and starts building possible models that satisfy both constraints. 
In addition to the given constraints, we have default rules to emulate
the CG principles: an analysis is true if no rule affects it,
and at least one analysis for each word is true---the notion of ``last'' is not applicable.
With these constraints, we get two solutions: \texttt{det, n} and \texttt{prn, v pri p3, v imp p2}. 
The interaction of the rules regarding \emph{la}  disambiguates the POS of \emph{casa} 
for free, and ordering of the rules doesn't matter. 
%Without more context or additional rules we cannot fully disambiguate the passage,
%but unlike with standard CG, we get only legal combinations in one solution.

@@

To replace the sequential order, we have experimented with different strategies:
\begin{itemize}
\item Maximise number of rules used---We discarded this strategy, because it was too strong: if a rule leads to a conflict in one case, it was not be applied anywhere.
\item Emulate sequential order---Enter clauses produced by each rule one by one, and assume the solver state reached so far is correct; if a conflict is introduced by new clauses, discard them and move on to next.
\item [\checkmark] Maximise number of rule applications---If there is a conflict, find the smallest number of rule applications to discard so that the conflict is solved.
\end{itemize}

%combined with the following
%\begin{itemize}
%\item Maximise number of true analyses---Without this heuristic, both solutions to our example %\emph{la casa} are as likely. With this heuristic,  the example would be disambiguated as \texttt{prn v}, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative.
% We want analyses to be true by default, but still disambiguate as much as possible.
%We have a rule to maximise the number of true analyses; then we take a solution that has the least. 
%\item Minimise number of true analyses---To combat the previous, we want each solution to contain as many true analyses as possible, to ensure that analyses untouched by rules are kept true. %After having multiple solutions which all ensure , we pick the one that disambiguates the most.
%\end{itemize}

@@

@@

SAT is NP-complete, whereas the standard implementations of CG are polynomial, 
but with advances in SAT solving techniques, 
the performance is much more feasible than in the previous works done in 90s--00s.
We used the open-source SAT solver MiniSat \cite{een04sat}.
%We do foreach sentence: foreach rule; 
%whereas VISL CG3 has foreach rule: foreach sentence.
%the size of the SAT problem is dependent of the sentence (???)

@@

\begin{table}
  \centering
  \begin{tabular}{|l|l|c|c|}
     \hline
	        \textbf{Text} & \textbf{\# words, rules} &  \textbf{SATCG} & \textbf{VISLCG3} \\ \hline
    \multirow{2}{*}{\textbf{Don Quijote}} & 186460 words  &  2m53.7s & 10.7s \\ 
		                                           & 261 rules & & \\ \hline
    \multirow{2}{*}{\textbf{Pride and}}  & 153386 words& 55.7s  & 4.8s \\ 
                         {\textbf{Prejudice}}                   & 99 rules & \\ \hline
  \end{tabular}
  \caption{Comparing execution times between SATCG and VISLCG3.}
  \label{table:time}
\end{table}

@@

\paragraph{Removing rules}
\begin{table}
  \centering
  \begin{tabular}{|l|c|c|}
     \hline
    \textbf{\# rules} & \textbf{SATCG} & \textbf{VISLCG3} \\
    \hline 
	19 &  83.15 & 83.33 \\ \hline  % 19 out of 23 preselected by authors
	18 &  83.19 & 83.04 \\ \hline  % 18 out of 19
	17 &  83.201 & 83.03 \\ \hline   % 17 out of 18/19, no diff
	16 &  83.197 & 83.030 \\ \hline  % 16 out of 17/18, no diff
	15 &  83.188 & 83.022\\ \hline  % 15 out of 16/17. Tie with 14 for SATCG, chose to retain the rule that gave better results for VISLCG3.
	14 &  83.180 & 83.005 \\ \hline  % Tie with 15 for SATCG.
	13 &  83.159 & 82.951 \\ \hline  % 13 out of 15 
	12 &  83.134 & 82.884\\ \hline  % 12 out of 14
	11 &  83.089 & 82.847\\ \hline  
	10 &  83.051 & 82.802\\ \hline  
	9  & 82.997 & 82.802 \\ \hline
	8  & 82.906 & 82.681 \\ \hline  % REMOVE:r_imp Imp IF (0 Subj) (-1* CnjSub) ; 
	7  & 82.743 & 82.594 \\ \hline  % SELECT:s_pro PrnIndep IF (1C VerbFin) ;
	6  & 82.556 & 82.398\\ \hline  %  REMOVE:r_v_v Verb IF (-1C Vblex) ;
	5  & 82.327 & 82.028 \\ \hline % SELECT:s_adv Adv IF (1 Cnj_Rel_End OR Prep OR Det)
	4  & 82.065 & 82.028 \\ \hline % REMOVE:r_vfin_prep VerbFin IF (0 Prep) (1 Inf) 
	3  & 81.387 & 81.233 \\ \hline % REMOVE:r_noun N    IF (-1 N OR Prn)
	2  & 80.23 & 79.97 \\ \hline % SELECT:s_pri_1 (pri p3) IF (0 (vblex imp p2 sg))
	1  & 76.69 & 76.62 \\ \hline 
	0  & 70.70 & 70.70 \\ \hline
  \end{tabular}
  \caption{Result of removing rules one by one, and how close to the gold standard they get.}
  \label{table:results_remove}
\end{table}
In this experiment, we took 19 rules written by the authors, which give 83.15 \% correct results with SATCG and 83.33 \% with VISLCG3.
Then we removed rules one by one; after getting a score for $n$ rules, we took the combination of $n-1$ rules for which SATCG gives the best score, and we ran the rule set with VISLCG3 for comparison. Table~\ref{table:results_remove} shows the percentages compared to the gold standard for both systems.
